### 2019-03-03 docs/notebooks/results.Rmd
fix up references in 'CNNs as visual system models' section of intro

### Convolutional neural networks and attentional mechanisms
Neural networks have in the past few years transitioned from being an academic area of artificial intellgience to being powerful tools for engineers working in areas such as image classification, speech-to-text, and automatic translation, to name but a few. In the 1980s, artificial neural networks trained with backpropagation gave rise to [Connectionist](https://en.wikipedia.org/wiki/Connectionism) models of cognition. Similar neural network models lie at the heart of the recent rise of deep learning, in large part because of advances in methods for training networks, and advances in technology such as GPUs that allow for parallelized training. When benchmarked with scalar metrics like accuracy on computer vision tasks such as object recognition, these deep neural networks appear capable of human or super-human levels of skill. The first major gains in accuracy were seen on image classification tasks such as ImageNet <cite data-cite="6027200/WGZWMWX7"></cite>, and the deep neural networks that made these advances were built with what are known as convolutional layers. As the name implies, these layers perform a convolution-like operation by multiplying filters of learned weights with the input image, producing as outputs feature maps that can be further transformed by succeeding layers. Convolutional neural networks (CNNs) represent the state of the art in image recognition. Many mechanisms proposed to extend the abilities of CNNs have been given the name "attention", and these attention mechanisms typically involve some sort of recurrence [@baMultipleObjectRecognition2014; @mnihRecurrentModelsVisual2014; @xuShowAttendTell2015]. These mechanisms are arguably analogous to what is called **overt** attention in psychology and neuroscience. In the context of vision, overt attention describes movement of the eyes or head so that the high-acuity fovea is placed on the focus of attention. In contrast, **covert** attention is any mechanism that occurs internally to select features relevant for a task while discarding irrelevant features. Covert attention can be considered one type of **dynamic routing** of information thought to occur in the brain. Models of dynamic routing, inspired by studies of visual attention [@bobierAttentionalRoutingCircuit2011; @bobierUnifyingMechanisticModel2014a; @olshausenMultiscaleDynamicRouting1995; @olshausenNeurobiologicalModelVisual1993] have also formed the basis of attention-like mechanisms in applied research on neural networks [@tangLearningGenerativeModels2014]. The function of CNNs has not typically been considered within the context of this work on covert visual attention.

### Convolutional neural networks as models of the visual system
Recently, many in neuroscience and cognitive science have returned to using neural network models to understand the brain, in part because of their resurgence. More specifically, many researchers have drawn parallels between the architecture of CNNs and the architecture of the visual system in the brain <cite data-cite="6027200/8ISULWGF"></cite>, in particular in humans and other primates where this system has been most thoroughly studied. Like CNNs, the visual system has a hierarchical structure, and is thought to function in part by performing transformations at each level of this hierarchy so that high-dimensional, low-level features are mapped into low-dimensional abstract representations. Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both <cite data-cite="6027200/6D4LKJD9"></cite><cite data-cite="6027200/U77Y74SI"></cite>. Representations can be easily extracted from CNNs because they are "transparent box" models--i.e. we have access to the "activity" of the individual units from which these models are built when we "show" the CNNs a stimulus. In the same studies, brain activity was obtained with electrophysiological or functional imaging methods. Such studies have found in general that representations are similar. (The comparison depends on representational similarity methods <cite data-cite="6027200/IQ2Y895S"></cite><cite data-cite="6027200/EQTU2P3Z"></cite>, but the concern of this study is not the validity of those methods.)

@@

## Results
In an initial set of simulations, we trained a CNN to "perform" the visual search task. We used the AlexNet architecture, which has previously been used in studies comparing CNNs to the visual system [@khaligh-razaviDeepSupervisedNot2014]. To train AlexNet, we followed a fine-tuning approach, where we re-used weights from training this CNN on the ImageNet dataset. During fine-tuning, the weights in the convolutional layers were frozen, but randomly initialized weights in the final fully-connected layers were update with a very small learning rate. The original softmax layer with 1000 outputs representing ImageNet classes was replaced with a 2-unit softmax layer, where one unit represented a response of "target present" and the other a response of "target absent". This approach is similar to that taken by a previous study [@poderCapacityLimitationsVisual2017]; we discuss differences in the details of fine-tuning in the methods below. To generate visual search stimuli that could be fed into a neural network, we used a simple Python package (<https://github.com/NickleDave/searchstims>) built on top of the [PyGame](https://www.pygame.org) library . The search displays were divided up into a 5x5 grid of cells, and items were plotted in a randomly chosen subset of cells (as described in <cite data-cite="6027200/QCQFYFDZ"></cite> and shown at <http://search.bwh.harvard.edu/new/data_set_files.html>). The library applied jitter to the items, so that two images could be different even if both images had the same set size and target present condition, and the items were in the exact same cells.  This is typically done with visual search stimuli used in psychophysics experiments, but an added benefit was that this allowed us to create a large set of unique training images (on the order of ten thousand) for each set size and condition, in effect resulting in a sort of data augmentation, as is often used when training neural networks.

>>>>

### Convolutional neural networks and attentional mechanisms
Neural networks have in the past few years transitioned from being an academic area of artificial intellgience to being powerful tools for engineers working in areas such as image classification, speech-to-text, and automatic translation, to name but a few. In the 1980s, artificial neural networks trained with backpropagation gave rise to [Connectionist](https://en.wikipedia.org/wiki/Connectionism) models of cognition. Similar neural network models lie at the heart of the recent rise of deep learning, in large part because of advances in methods for training networks, and advances in technology such as GPUs that allow for parallelized training. When benchmarked with scalar metrics like accuracy on computer vision tasks such as object recognition, these deep neural networks appear capable of human or super-human levels of skill. The first major gains in accuracy were seen on image classification tasks such as ImageNet [@krizhevskyImagenetClassificationDeep2012; @simonyanVeryDeepConvolutional2014], and the deep neural networks that made these advances were built with what are known as convolutional layers. As the name implies, these layers perform a convolution-like operation by multiplying filters of learned weights with the input image, producing as outputs feature maps that can be further transformed by succeeding layers. Convolutional neural networks (CNNs) represent the state of the art in image recognition. Many mechanisms proposed to extend the abilities of CNNs have been given the name "attention", and these attention mechanisms typically involve some sort of recurrence [@baMultipleObjectRecognition2014; @mnihRecurrentModelsVisual2014; @xuShowAttendTell2015]. These mechanisms are arguably analogous to what is called **overt** attention in psychology and neuroscience. In the context of vision, overt attention describes movement of the eyes or head so that the high-acuity fovea is placed on the focus of attention. In contrast, **covert** attention is any mechanism that occurs internally to select features relevant for a task while discarding irrelevant features. Covert attention can be considered one type of **dynamic routing** of information thought to occur in the brain. Models of dynamic routing, inspired by studies of visual attention [@bobierAttentionalRoutingCircuit2011; @bobierUnifyingMechanisticModel2014a; @olshausenMultiscaleDynamicRouting1995; @olshausenNeurobiologicalModelVisual1993] have also formed the basis of attention-like mechanisms in applied research on neural networks [@tangLearningGenerativeModels2014]. The function of CNNs has not typically been considered within the context of this work on covert visual attention.

### Convolutional neural networks as models of the visual system
Recently, many in neuroscience and cognitive science have returned to using neural network models to understand the brain, in part because of their resurgence. More specifically, many researchers have drawn parallels between the architecture of CNNs and the architecture of the visual system in the brain [@kriegeskorteDeepNeuralNetworks2015], in particular in humans and other primates where this system has been most thoroughly studied. Like CNNs, the visual system has a hierarchical structure, and is thought to function in part by performing transformations at each level of this hierarchy so that high-dimensional, low-level features are mapped into low-dimensional abstract representations. Several recent studies in the realm of cognitive neuroscience have measured representations in more directly tested the similarities between CNNs and the visual system, by comparing representations in both [@khaligh-razaviDeepSupervisedNot2014; @yaminsPerformanceoptimizedHierarchicalModels2014; @yaminsUsingGoaldrivenDeep2016]. Representations can be easily extracted from CNNs because they are "transparent box" models--i.e. we have access to the "activity" of the individual units from which these models are built when we "show" the CNNs a stimulus. In the same studies, brain activity was obtained with electrophysiological or functional imaging methods. Such studies have found in general that representations are similar. (Of course this comparison depends on methods for measuring representational similarity [@bobadilla-suarezMeasuresNeuralSimilarity2018; @diedrichsenRepresentationalModelsCommon2017], but this study does not address those methods.)

@@

## Results
In an initial set of simulations, we trained a CNN to "perform" the visual search task. We used the AlexNet architecture, which has previously been used in studies comparing CNNs to the visual system [@khaligh-razaviDeepSupervisedNot2014]. To train AlexNet, we followed a fine-tuning approach, where we re-used weights from training this CNN on the ImageNet dataset. During fine-tuning, the weights in the convolutional layers were frozen, but randomly initialized weights in the final fully-connected layers were update with a very small learning rate. The original softmax layer with 1000 outputs representing ImageNet classes was replaced with a 2-unit softmax layer, where one unit represented a response of "target present" and the other a response of "target absent". This approach is similar to that taken by a previous study [@poderCapacityLimitationsVisual2017]; we discuss differences in the details of fine-tuning in the methods below. To generate visual search stimuli that could be fed into a neural network, we used a simple Python package (<https://github.com/NickleDave/searchstims>) built on top of the [PyGame](https://www.pygame.org) library . The search displays were divided up into a 5x5 grid of cells, and items were plotted in a randomly chosen subset of cells (as described in [@wolfeReactionTimeDistributions2010] and shown at <http://search.bwh.harvard.edu/new/data_set_files.html>). The library applied jitter to the items, so that two images could be different even if both images had the same set size and target present condition, and the items were in the exact same cells.  This is typically done with visual search stimuli used in psychophysics experiments, but an added benefit was that this allowed us to create a large set of unique training images (on the order of ten thousand) for each set size and condition, in effect resulting in a sort of data augmentation, as is often used when training neural networks.
